Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Shanghai version \YellowPaperVersionNumber}}

### 2019-11-16 Paper.tex
Fix syntax error.
Prefer MUST to SHOULD for requirements

\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
\author{
    Dr. Gavin Wood\\
    Founder, Ethereum \& Parity\\
    gavin@parity.io
}
\begin{document}

@@

With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap. Technology-rooted movements like Bitcoin have demonstrated through the power of the default, consensus mechanisms, and voluntary respect of the social contract, that it is possible to use the internet to make a decentralised value-transfer system that can be shared across the world and virtually free to use. This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine. Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications albeit rather simplistic ones.

@@

%E language?
For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.

@@

Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. It is this final state which we accept as the canonical ``version'' of the world of Ethereum. The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible. Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes. Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere. A valid state transition is one which comes about through a transaction. Formally:
\begin{equation}
\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
\end{equation}

@@

Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference. Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big). They also punctuate the transaction series with incentives for nodes to \textit{mine}. This incentivisation takes place as a state-transition function, adding value to a nominated account.

Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block. It is achieved thanks to a cryptographically secure proof. This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...), ...) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.

@@

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:
\par
\begin{center}
\begin{tabular}{rl}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

@@

The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}. This process is described in detail in section \ref{ch:ghost}.

Sometimes, a path follows a new protocol from a particular height.  This document describes one version of the protocol.  In order to follow back the history of a path, one must reference multiple versions of this document.

@@

The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}). Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}). The trie requires a simple database backend that maintains a mapping of byte arrays to byte arrays; we name this underlying database the state database. This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state. Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly. Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.

@@

\begin{description}
\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account. For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values. The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval. This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
\end{description}

Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
\begin{equation}
\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
\end{equation}

The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
\begin{equation}
L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
\end{equation}

@@

Thus we may define a world-state collapse function $L_{S}$:
\begin{equation}
L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
\end{equation}

This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
\end{equation}

@@

A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum. While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents. \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}. There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation'). Both types specify a number of common fields:

\begin{description}
\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction. This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}. This is expanded in Appendix \ref{app:signing}.
\end{description}

Additionally, a contract creation transaction contains:

@@

\begin{equation}
L_{T}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
\end{equation}

@@

@@

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}). The block header contains several pieces of information:

@@

\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
\begin{equation}
R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
\end{equation}

>>>>

\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Shanghai version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Shanghai version \YellowPaperVersionNumber}}}
\author{
    Dr. Gavin Wood\\
    Founder, Ethereum \& Parity\\
    gavin@parity.io
}
\begin{document}

@@

With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap. Technology-rooted movements like Bitcoin have demonstrated through the power of the default, consensus mechanisms, and voluntary respect of the social contract, that it is possible to use the internet to make a decentralised value-transfer system that can be shared across the world and virtually free to use. This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine. Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications, albeit rather simplistic ones.

@@

%E language?
For a list of terms used in this paper, refer to Appendix~\ref{ch:Terminology}.

@@

Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some current state. It is this current state which we accept as the canonical ``version'' of the world of Ethereum. The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible. Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes. Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere. A valid state transition is one which comes about through a transaction. Formally:
\begin{equation}
\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
\end{equation}

@@

Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference. Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...), ...) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...
\end{eqnarray}

Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function for transactions\footnote{Note that since the Shanghai fork, blocks also needs to process \textit{withdrawal operations} in order to reach their final state. Withdrawal operations are defined in sub-section \ref{subsec:The_Withdrawal}, and block final state is discussed in greater detail in section \ref{ch:finalisation}.}.

@@

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:
\par
\begin{center}
\begin{tabular}{rl}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^9$ & Gwei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

@@

Since the \textit{Paris} hard fork, reaching consensus on new blocks is managed by a protocol called the \textit{Beacon Chain}.
It is known as the \textit{consensus layer} of Ethereum, and it defines the rules for determining the canonical history of Ethereum blocks.
This document describes the \textit{execution layer} of Ethereum.
The execution layer defines the rules for interacting with and updating the state of the Ethereum virtual machine.
The consensus layer is described in greater detail in the \href{https://github.com/ethereum/consensus-specs}{consensus specifications}.
How the consensus layer is used to determine the canonical state of Ethereum is discussed in section \ref{ch:blocktree_to_blockchain}.

There are many versions of Ethereum, as the protocol has undergone a number of updates.
These updates can be specified to occur:
\begin{itemize}
  \item at a particular block number in the case of pre-\textit{Paris} updates,
  \item after reaching a \textit{terminal total difficulty} in the case of the \textit{Paris} update, or
  \item at a particular block timestamp in the case of post-\textit{Paris} updates.
\end{itemize}
This document describes the \textit{Shanghai} version.

In order to follow back the history of a path, one must reference multiple versions of this document.
Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris, Shanghai, and every upcoming forks activated at a given block number (e.g. 15,537,394 for Paris), the trigger was not the block number, but rather reaching a specified \textit{timestamp} (or \textit{total difficulty} for Paris). The trigger for the \textit{Paris} and subsequent hard fork are discussed in greater detail in section \ref{ch:pos_transition}.}
\par
\begin{center}
\begin{tabular}{lr}
\toprule
Name & First Block Number \\
\midrule
$F_{\mathrm{Homestead}}$         &  1150000 \\
$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
$F_{\mathrm{Byzantium}}$         &  4370000 \\
$F_{\mathrm{Constantinople}}$    &  7280000 \\
$F_{\mathrm{Petersburg}}$        &  7280000 \\
$F_{\mathrm{Istanbul}}$          &  9069000 \\
$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
$F_{\mathrm{Berlin}}$            & 12244000 \\
$F_{\mathrm{London}}$            & 12965000 \\
$F_{\mathrm{Arrow Glacier}}$     & 13773000 \\
$F_{\mathrm{Gray Glacier}}$      & 15050000 \\
$F_{\mathrm{Paris}}$             & 15537394 \\
$F_{\mathrm{Shanghai}}$          & 17034870 \\
\bottomrule
\end{tabular}
\end{center}
\par

Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
For the Ethereum main network
\begin{equation}
  \linkdest{chain_id}
  \beta = 1
\end{equation}

@@

The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix~\ref{app:rlp}). Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}). The trie requires a simple database backend that maintains a mapping of byte arrays to byte arrays; we name this underlying database the state database. This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state. Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly. Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.

@@

\begin{description}
\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account. For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values. The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call. All such code fragments are contained in the state database under their corresponding hashes for later retrieval. This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
\end{description}

Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
\begin{equation}
\texttt{TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
\end{equation}

The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
\begin{equation}
L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
\end{equation}

@@

Thus we may define a world-state collapse function $L_{\mathrm{S}}$:
\begin{equation}
L_{\mathrm{S}}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
\end{equation}

This function, $L_{\mathrm{S}}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
\end{equation}

@@

A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
The sender of a transaction cannot be a contract.
While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably,
such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
\eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
EIP-2718 by \cite{EIP-2718} introduced the notion of different transaction types.
As of the \textit{London} version of the protocol, there are three transaction types: 0 (legacy), 1 (EIP-2930 by \cite{EIP-2930}), and 2 (EIP-1559 by \cite{EIP-1559}).
Further, there are two subtypes of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
All transaction types specify a number of common fields:

\begin{description}
\item[type]\linkdest{tx_type}{} EIP-2718 transaction type; formally $T_{\mathrm{x}}$.
\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction. This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
\item[r, s]\linkdest{T__r_T__s}{} Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}. This is expanded in Appendix \ref{app:signing}.
\end{description}

EIP-2930 (type 1) and EIP-1559 (type 2) transactions also have:
\begin{description}
\item[accessList]\linkdest{tx_access_list}{} List of access entries to warm up; formally $T_{\mathbf{A}}$.
\linkdest{access_list_entry}{}Each access list entry $E$ is a tuple of an account address and a list of storage keys: $E \equiv (E_{\mathrm{a}}, E_{\mathbf{s}})$. 
\item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
\item[yParity]\linkdest{tx_y_parity}{} Signature Y parity; formally $T_{\mathrm{y}}$.
\end{description}

Legacy transactions do not have an \textbf{accessList} ($T_{\mathbf{A}}=()$), while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
\begin{description}
\item[w]\linkdest{T__w}{} A scalar value encoding Y parity and possibly chain ID; formally $T_{\mathrm{w}}$.
$T_{\mathrm{w}} = 27 + T_{\mathrm{y}}$ or $T_{\mathrm{w}} = 2\beta + 35 + T_{\mathrm{y}}$ (see EIP-155 by \cite{EIP-155}).
\end{description}

There are differences in how one's acceptable gas price is specified in type 2 transactions versus type 0 and type 1 transactions. Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee}\footnote{The \textit{priority fee} is discussed in greater detail in sections \ref{ch:payment} and \ref{ch:transactions}.} that is paid. Type 2 transactions have the following two fields related to gas:

\begin{description}
  \item[maxFeePerGas]\linkdest{max_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{m}}$.
  \item[maxPriorityFeePerGas]\linkdest{max_priority_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid to the block's fee recipient as an incentive to include the transaction; formally $T_{\mathrm{f}}$.
\end{description}

In contrast, type 0 and type 1 transactions specify gas price as a single value:

\begin{description}
  \item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.\footnote{Type 0 and type 1 transactions will get the same gas price behavior as a type 2 transaction with $T_{\mathrm{m}}$ and $T_{\mathrm{f}}$ set to the value of $T_{\mathrm{p}}$.}
\end{description}

Additionally, a contract creation transaction (regardless of transaction type) contains:

@@

\begin{equation}
\linkdest{L_transaction} L_{\mathrm{T}}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 0 \\
(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{f}}, T_{\mathrm{m}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
\end{cases}
\end{equation}
where
\begin{equation}
\mathbf{p} \equiv \begin{cases}
T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
T_{\mathbf{d}} & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{A}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_{\mathrm{x}} \in \{0, 1, 2\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B} & \wedge \\
T_{\mathrm{m}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{f}} \in \mathbb{N}_{256}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
\end{equation}

@@

\subsection{The Withdrawal}\linkdest{withdrawal}\label{subsec:The_Withdrawal}
A withdrawal (formally, \textit{W}) is a tuple of data describing a consensus layer validator's withdrawal of some amount of its staked Ether.
A withdrawal is created and validated in the consensus layer of the blockchain and then pushed to the execution layer.
A withdrawal is composed of the following fields:

\begin{description}
\item[globalIndex] zero based incrementing withdrawal index that acts as a unique identifier for this withdrawal; formally $W_{\mathrm{g}}$.
\item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
\item[recipient] the 20-byte address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
\item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
\end{description}

Withdrawal serialisation is defined as:
\begin{equation}
\linkdest{L_Withdrawal} L_{\mathrm{W}}(W) \equiv 
(W_{\mathrm{g}}, W_{\mathrm{v}}, W_{\mathrm{r}}, T_{\mathrm{a}})
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values except for $W_\mathbf{r}$ which is a 20-byte address:
\begin{equation}
\begin{array}[t]{lclc}
W_{\mathrm{g}} \in \mathbb{N}_{64} & \wedge & W_{\mathrm{v}} \in \mathbb{N}_{64} & \wedge \\
W_{\mathrm{r}} \in \mathbb{B}_{20} & \wedge & W_{\mathrm{a}} \in \mathbb{N}_{64}
\end{array}
\end{equation}

@@

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions,
$\mathbf{T}$,\hypertarget{ommerheaders}{} a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}),
and, since the \textit{Shanghai} hard fork, $\mathbf{W}$, a collection of validator's withdrawal pushed by the consensus layer. The block header contains several pieces of information:

@@

\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of five items comprising:
the type of the transaction, $R_{\mathrm{x}}$,
the status code of the transaction, $R_{\mathrm{z}}$,
the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$,
the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$}:
\begin{equation}
R \equiv (R_{\mathrm{x}}, R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
\end{equation}

$R_{\mathrm{x}}$ is equal to the \hyperlink{tx_type}{type} of the corresponding transaction.
